---
title: "Network interference estimators"
author: "Vanessa McNealis"
date: "2024-01-18"
output: pdf_document
---

Load dependencies and source script

```{r dependencies, message = FALSE, warning = FALSE}
library(boot)
library(tidyverse)
library(igraph)
library(data.table)
library(pbapply)
library(lme4)
library(statnet)
library(intergraph)
library(statmod)
library(fastGHQuad)
library(geex)
library(here)
source(url("https://github.com/vanessamcnealis/Network-Interference-estimators/blob/main/Estimating_functions.R?raw=TRUE"))
```

## Data preparation

The data at hand \texttt{dat} consists of one simulated dataset used for the simulation
studies presented in the paper. The data generating scheme for the assigned treatment 
$Z$, the outcome $Y$, and the covariates $X_1$, $X_2$ and $X_3$ is described in the
manuscript. 

```{r prep}
dat  <- read.csv("https://github.com/vanessamcnealis/Network-Interference-estimators/blob/main/dat.csv?raw=TRUE")
load(url("https://github.com/vanessamcnealis/Network-Interference-estimators/blob/main/network.RData?raw=TRUE"))
```

### Network visualization

```{r plot}
plot(network, vertex.size=1, vertex.label=NA, vertex.color="cadetblue3", 
     vertex.frame.color="cadetblue3")
```

### Prepare the data for modeling

```{r covariates}
# Total number of nodes
N <- nrow(dat)

#Number of components
m <- length(unique(dat$component.id))

# Adjacency matrix associated with the graph
A <- get.adjacency(network)

# Create a variable indicating the degree of a node
dat$degree <- as.numeric(A %*% rep(1, N))

# Create a variable indicating the number of treated neighbours for each node
dat$k_treated <- 
    as.numeric(A %*% matrix(dat$Z, ncol = 1, nrow = nrow(dat)))

# Create a variable indicated the proportion of treated neighbours for each node
dat$prop_treated <- ifelse(dat$degree == 0, 0, dat$k_treated/dat$degree)

# Change this vector for different alpha values
alpha <- c(0.2, 0.5, 0.8)

```

## Modeling

In this section, we will apply the IPW, REG, DR-BC, and DR-WLS estimators for the 
average potential outcomes and associated causal contrasts. The average potential 
outcome for a unit depends on an individual exposure/treatment $z$ and 
treatment coverage $\alpha \in (0,1)$, where the treatment coverage represents the 
counterfactual probability that first-order neighbours receive the 
exposure/treatment. For each estimation method, the R programs will output the 
\textbf{point estimate and the estimated variance for average potential
outcomes $\mu_{0\alpha}, \mu_{1\alpha}, \mu_{\alpha}$} under three different treatment coverages (i.e., $\alpha \in \{0.2, 0.5, 0.8\}$), 
respectively. Also, the R program will output the point estimation and estimated 
variance of four causal effects: Direct (DE), Indirect (IE), Total (TE), and 
Overall effect (OE).

Equations for calculating the four causal effect estimates are:
\begin{itemize}
\item $\hat{DE}(\alpha) =  \hat{\mu}_{1\alpha} - \hat{\mu}_{0\alpha}$
\item $\hat{IE}(\alpha_0, \alpha_1) =  \hat{\mu}_{0\alpha_1} - \hat{\mu}_{0\alpha_0}$
\item $\hat{TE}(\alpha_0, \alpha_1) =  \hat{\mu}_{1\alpha_1} - \hat{\mu}_{0\alpha_0}$
\item $\hat{OE}(\alpha_0, \alpha_1) =  \hat{\mu}_{1\alpha_1} - \hat{\mu}_{0\alpha_0},$
\end{itemize}
where $\alpha_0$ and $\alpha_1$ represent distinct treatment coverages with 
$\alpha_0 \neq \alpha_1$. If the vector supplied is $\alpha = (0.2, 0.5, 0.8)$,
the program will treat values in sequential order and compare
\begin{itemize}
\item $\alpha_1 = 0.5$ vs. $\alpha_0 = 0.2$
\item $\alpha_1 = 0.8$ vs. $\alpha_0 = 0.2$
\item $\alpha_1 = 0.8$ vs. $\alpha_0 = 0.5$.
\end{itemize}

We first specify the models for the outcome and treatment propensity (note that
these correspond to the oracle models). In the model for the treatment propensity,
we include a random effect term for the component/subgraph. Note that for DR-WLS, the outcome formula
supplied should not include terms for the treatment assignment as this method
stratifies the sample based on the treatment indicator. 

```{r formulas}

propensity_formula <- Z ~ abs(X1) + I(X2*abs(X1)) +X3+ (1|component.id)
outcome_formula <- Y ~ Z + prop_treated + prop_treated*Z + abs(X1) + X2 + 
  abs(X1)*X2
outcome_formula_drwls <- Y ~ prop_treated + abs(X1) + X2 + abs(X1)*X2

```


### Inverse probability-of-treatment weighting (IPW) estimates

```{r ipw}

ipw_estimates <- ipw_estimation(data=dat, 
                                propensity_formula = propensity_formula, 
                                alpha = alpha)
ipw_estimates

```

In the output above,
\begin{itemize}
\item \textbf{Section [[1]]} displays the point estimates, estimated variances, and 95\% Wald-type confidence intervals for the average potential outcomes under three different treatment coverages (i.e., $\alpha = 0.2, 0.5, 0.8$) and each individual exposure (i.e. $z=0,1$ , and marginal). The marginal average potential outcome is the average potential outcomes for a particular treatment coverage, regardless of individual exposure status.
\item \textbf{Section [[2]]} displays the point estimates (estimand = “Direct effect”, “Indirect effect”, “Total effect”, “Overall effect” in the output), estimated variances, and 95\% Wald-type confidence intervals for four causal contrasts: direct, indirect, total and overall effects, corresponding to two distinct treatment coverages $\alpha_0$ and $\alpha_1$.
\end{itemize}

### Outcome regression (REG) estimates

```{r reg}

reg_estimates <- reg_estimation(data=dat, 
                                outcome_formula = outcome_formula, 
                                alpha = alpha)
reg_estimates

```

See the section \textbf{Inverse probability-of-treatment weighting (IPW) estimates} for 
an interpretation of the output. 

### Regression estimation with residual bias correction (DR-BC) estimates

```{r drbc}

drbc_estimates <- drbc_estimation(data=dat, 
                                  propensity_formula=propensity_formula,
                                  outcome_formula = outcome_formula, 
                                  alpha = alpha)
drbc_estimates

```

See the section \textbf{Inverse probability-of-treatment weighting (IPW) estimates} for 
an interpretation of the output. 

### Regression estimation with inverse-propensity weighted coefficients (DR-WLS) estimates

```{r drwls}

drwls_estimates <- drwls_estimation(data=dat, 
                                    propensity_formula=propensity_formula,
                                    outcome_formula = outcome_formula_drwls, 
                                    alpha = alpha)
drwls_estimates

```

See the section \textbf{Inverse probability-of-treatment weighting (IPW) estimates} for 
an interpretation of the output. 


